---
title: "PA1_template"
output: html_document
---
Read the data given in the assignment into activityData, we take into consideration only the complete cases, and discard NA values and store the result in ComplActivity.

Calculate the data frames for part 1 & 2 of the assignments which will help us build the plots.

```{r simulatedata, echo=TRUE}
activityData <- read.csv("activity.csv")
library(dplyr)
library(ggplot2)
ComplActivity <- activityData[complete.cases(activityData),]
activity_byday <- group_by(ComplActivity, date)

steps_sum <- summarize(activity_byday, sum(steps))
colnames(steps_sum) <- c("day", "Total_steps")
activity_byinterval <- group_by(ComplActivity, interval)

avgSteps <- summarize(activity_byinterval, mean(steps))
colnames(avgSteps) <- c("interval", "average_steps")


```
## Part 1 plot
Plot a histogram of the total number of steps taken per day and calculate Mean & Median of the total number of steps taken per day.

```{r histogramPlot, fig.height=4, echo = TRUE}
qplot(Total_steps,data = steps_sum)
meanSteps <- mean(steps_sum$Total_steps)
medianSteps <- median(steps_sum$Total_steps)

```

Now we got our mean which is `r meanSteps` and median which is `r medianSteps` of the total number of steps taken per day.

## Part 2 plot

Make a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis).
Calculate the  5-minute interval, on average across all the days in the dataset, contains the maximum number of steps.


``` {r timePlot, fig.height = 4, echo=TRUE}
plot(avgSteps$interval,avgSteps$average_steps, typ="l", xlab="5 min interval", ylab = "Average steps")
max <- avgSteps[which.max(avgSteps$average_steps),]
MaxInt <- max$interval
```

The 5 min interval, on the average across all the the days in the dataset, contains the maximum number of steps is `r MaxInt`

## Part 3 
The na values in the data set have been replaced by 200. 

``` {r part3, echo=TRUE}
activityData <- read.csv("activity.csv")
NaValues <- length(which(is.na(activityData$steps)))
activityData[["steps"]][is.na(activityData[["steps"]])] <- 200
NewActivityData <- activityData
NewAct_byday <- group_by(NewActivityData, date)
NewStepsSum <- summarize(NewAct_byday, sum(steps))
colnames(NewStepsSum) <- c("day", "Total_steps")
qplot(Total_steps, data= NewStepsSum)
NewMeanSteps <- mean(NewStepsSum$Total_steps)
NewMedianSteps <- median(NewStepsSum$Total_steps)

```

Total number of missing values in the dataset is `r NaValues`. The new mean is `r NewMeanSteps` and new median is `r NewMedianSteps`. Which seem to be very different from the part 1.

## Part 4

```{r part4, echo=TRUE}
NewActivityData4 <- mutate(NewActivityData, day = weekdays(as.Date(date)))

```




